\Large Postdoctoral Position Research Proposal: \\ The Role of Disk Volatile Chemistry and Dynamics in Shaping the Compositions of Nascent Planets\vspace{-2ex}

### 2017-06-15 Postdoc Applications/proposal_draft.tex
upd

%\usepackage{pdfsync}
%\usepackage{subeqnarray}
\usepackage[top=0.9in, bottom=0.6in, left=1in, right=1in]{geometry}
\usepackage{natbib}
\usepackage{color}
\usepackage{graphicx}
\usepackage{fancyhdr}
\usepackage[T1]{fontenc}
\usepackage{titling}
\usepackage{sectsty}
\usepackage{sidecap}
\usepackage{placeins}
\usepackage{indentfirst}
\usepackage{wrapfig}
\usepackage[font={footnotesize}]{caption}
\setlength{\droptitle}{-7em}
%\setlength{\abovecaptionskip}{-2.5ex}
\setlength{\belowcaptionskip}{-1ex}
%\pagenumbering{gobble}
\pagestyle{fancy}
\lhead{Ana-Maria Piso}
\rhead{Postdoctoral Position Research Proposal}
\date{}
%\rhead{\thepage}

@@

\title{\Large Postdoctoral Position Research Proposal: \\
The Role of Disk Volatile Chemistry and Dynamics in Shaping the Compositions of Nascent Planets\vspace{-2ex}}
\author{Ana-Maria Piso \vspace{-2ex}}

@@

Chemical and dynamical processes in a protoplanetary disk affect the disk structure and composition, and thus the composition of nascent planets. As shown in Figure 1, the timescales for various disk chemical and dynamical processes may be comparable at least in some parts of the disk. It follows that these two effects cannot be decoupled, and that the chemo-dynamical evolution of the disk has to be studied simultaneously and self-consistently. Through \textbf{analytical and numerical calculations}, I will first explore a \textbf{range of dynamical processes that may affect the distribution of volatiles in disks}, expanding and generalizing the framework I developed during my dissertation research $[7]$. Such effects include particle growth and fragmentation, as well as accretion rate and stellar luminosity evolution.

\begin{figure}[h!]
\centering
\includegraphics[width=\textwidth]{chemical_timescales}
\vspace{-0.3in}
\caption{The distribution of characteristic chemical and dynamical timescales $\tau$ in disks as a function of semi-major axis and height. The numbered curves are $\log_{10} (\tau)$ in million years (from $[11]$).}
\label{fig:chemical}
\end{figure}

@@

The complex processes outlined in part 1 will directly impact the composition and dynamics of forming gas giants, as the latter are born and evolve simultaneously with the disk. 
\begin{wrapfigure}{r}{0.42\textwidth}
\vspace{-0.4in}
\begin{center}
\includegraphics[width=0.42\textwidth]{HLTau_nrao}
\end{center}
%\vspace{-0.5in}
\caption{ALMA continuum image of the disk around HL Tau, which shows evidence of a gas gap in the disk, most likely created due to planetary migration (from $[8]$).}
\vspace{-0.25in}
\end{wrapfigure}
The chemical evolution of the gas disk molecular abundances, as well as disk dynamics, determine how a young planet's atmospheric composition changes in time and with the planet's location, and therefore the final chemical structure of mature planets. \textbf{It is thus clear that disks and planets are deeply connected, and that this relation needs to be explored}. %I will add planetary embryos in the code developed in step 1, and

Giant planets can migrate through the disk while still accumulating gas. Figure 2 shows an exciting ALMA observation of a planetary gap in the disk around HL Tau $[8]$. This will change a planet's atmospheric composition since the disk chemical abundances are different at different disk locations. Additionally, giant planets may still accumulate planetesimals while accreting nebular gas $[9]$. The final composition of a planet's atmosphere will thus depend on how much gas and solids are accreted in this stage. \textbf{I will add planet dynamical effects such as migration and planetesimal accretion in the chemical and dynamical model developed in part 1, and quantify how these processes affect the chemical composition of gas giant envelopes}.

@@

The results from parts 1 and 2 will feed into a \textbf{large planet synthesis model}, in which I will use a grid of different \textbf{initial disk and planetary embryo conditions}. I will develop my own numerical program instead of using already existing codes of planet population synthesis (e.g., $[10]$ and references therein), since the latter involve many complex physical processes, some of which might not be necessary or relevant for my own calculations. To minimize computational expenses, I will build a program from bottom-up that includes select physical and chemical processes based on the local simulations from steps 1 and 2. This will allow me to predict the \textbf{range of planet compositions that can form under reasonable disk conditions}, as well as \textbf{constrain a planet's formation location based on its chemical composition}. Comparing my results with current observations of atmospheric spectra (e.g., Figure 3 from $[4]$), and more importantly future JWST observations, will lead to great scientific strides in understanding the complex connection between protoplanetary disks and the formation, evolution and composition of exoplanets.

@@

\textbf{MIT is the ideal place for me to pursue my postdoctoral work.} Its research facilities and academic excellence are unparalleled. I would love the opportunity to work with \textbf{Prof. Hilke Schlichting}, an expert in planet formation theory and dynamics --- such a collaboration would be instrumental in achieving my research goals during my postdoctoral tenure and afterwards. I would also like to collaborate with \textbf{Prof. Sara Seager}, a co-investigator of the TESS mission and a leading theorist in atmospheric chemistry, as well as with members of her group, who specialize in a broad range of exoplanet theory and computational topics. \textbf{Dr. Margaret Pan}, an excellent theorist in planetary dynamics and protoplanetary disks, will join the MIT EAPS department in early 2016 --- it would be a pleasure to work with her as well. Additionally, the MIT Physics department hosts leaders in detecting and characterizing worlds outside the Solar system, such as \textbf{Prof. Joshua Winn}, which presents great prospects in connecting my theoretical research work with observations. \\

@@

%\section*{References}
\footnotesize
%\begin{thebibliography} {9}
\noindent $[1]$ Batalha, N. M. 2014, Proceedings of the National Academy of Science, 111, 12647 \\
$[2]$ Lissauer J. J., Dawson R. I., Tremaine S., 2014, Nature, 513, 336 \\
$[3]$ Debes, J. H., Jang-Condell, H., Weinberger, A. J., Roberge, A., \& Schneider, G. 2013, ApJ, 771, 45 \\
$[4]$ Kreidberg, L., Line, M. R., Bean, J. L., et al. 2015, ApJ, 814, 66 \\
$[5]$ Piso, A.-M. A., \& Youdin, A. N. 2014, ApJ, 786, 21 \\
$[6]$ Piso, A.-M. A., Youdin, A. N., \& Murray-Clay, R. A. 2015, ApJ, 800, 82 \\
$[7]$ Piso, A.-M. A., \"Oberg, K. I. , Birnstiel, T., \& Murray-Clay, R. A. 2015, ApJ, 815, 109 \\
$[8]$ ALMA Partnership et al., 2015, ApJ, 808, L3 \\
$[9]$ \"Oberg, K. I., Murray-Clay, R., \& Bergin, E. A. 2011, ApJ, 743, L16 \\
$[10]$ Ida, S., Lin, D. N. C., \& Nagasawa, M. 2013, ApJ, 775, 42 \\
$[11]$ Semenov, D., \& Wiebe, D. 2011, ApJS, 196, 25
%\end{thebibliography}

>>>>

%\usepackage{pdfsync}
%\usepackage{subeqnarray}
\usepackage[top=0.9in, bottom=0.6in, left=1in, right=1in]{geometry}
\usepackage{natbib}
\usepackage{color}
\usepackage{graphicx}
\usepackage{fancyhdr}
\usepackage[T1]{fontenc}
\usepackage{titling}
\usepackage{sectsty}
\usepackage{sidecap}
\usepackage{placeins}
\usepackage{indentfirst}
\usepackage{wrapfig}
\usepackage[font={footnotesize}]{caption}
\setlength{\droptitle}{-7em}
%\setlength{\abovecaptionskip}{-2.5ex}
\setlength{\belowcaptionskip}{-1ex}
%\pagenumbering{gobble}
\pagestyle{fancy}
\lhead{Dr. Ana-Maria Piso}
\rhead{Postdoctoral Position Research Proposal}
\date{}
%\rhead{\thepage}

@@

\title{\Large Postdoctoral Position Research Proposal: \\
The Role of Disk Volatile Chemistry and Dynamics in Shaping the Compositions of Nascent Planets\vspace{-2ex}}
\author{Dr. Ana-Maria Piso \vspace{-2ex}}

@@

Chemical and dynamical processes in a protoplanetary disk affect the disk structure and composition, and thus the composition of nascent planets. As shown in Figure 1, the timescales for various disk chemical and dynamical processes may be comparable at least in some parts of the disk. It follows that these two effects cannot be decoupled, and that the chemo-dynamical evolution of the disk has to be studied simultaneously and self-consistently. Through \textbf{analytical and numerical calculations}, I will first explore a \textbf{range of dynamical processes that may affect the distribution of volatiles in disks}, expanding and generalizing the framework I developed during my dissertation research ($[7]$, $[8]$). Such effects include particle growth and fragmentation, as well as accretion rate and stellar luminosity evolution.

\begin{figure}[h!]
\centering
\includegraphics[width=\textwidth]{chemical_timescales}
\vspace{-0.3in}
\caption{The distribution of characteristic chemical and dynamical timescales $\tau$ in disks as a function of semi-major axis and height. The numbered curves are $\log_{10} (\tau)$ in million years (from $[12]$).}
\label{fig:chemical}
\end{figure}

@@

The complex processes outlined in part 1 will directly impact the composition and dynamics of forming gas giants, as the latter are born and evolve simultaneously with the disk. 
\begin{wrapfigure}{r}{0.42\textwidth}
\vspace{-0.4in}
\begin{center}
\includegraphics[width=0.42\textwidth]{HLTau_nrao}
\end{center}
%\vspace{-0.5in}
\caption{ALMA continuum image of the disk around HL Tau, which shows evidence of a gas gap in the disk, most likely created due to planetary migration (from $[9]$).}
\vspace{-0.25in}
\end{wrapfigure}
The chemical evolution of the gas disk molecular abundances, as well as disk dynamics, determine how a young planet's atmospheric composition changes in time and with the planet's location, and therefore the final chemical structure of mature planets. \textbf{It is thus clear that disks and planets are deeply connected, and that this relation needs to be explored}. %I will add planetary embryos in the code developed in step 1, and

Giant planets can migrate through the disk while still accumulating gas. Figure 2 shows an exciting ALMA observation of a planetary gap in the disk around HL Tau $[8]$. This will change a planet's atmospheric composition since the disk chemical abundances are different at different disk locations. Additionally, giant planets may still accumulate planetesimals while accreting nebular gas $[10]$. The final composition of a planet's atmosphere will thus depend on how much gas and solids are accreted in this stage. \textbf{I will add planet dynamical effects such as migration and planetesimal accretion in the chemical and dynamical model developed in part 1, and quantify how these processes affect the chemical composition of gas giant envelopes}.

@@

The results from parts 1 and 2 will feed into a \textbf{large planet synthesis model}, in which I will use a grid of different \textbf{initial disk and planetary embryo conditions}. I will develop my own numerical program instead of using already existing codes of planet population synthesis (e.g., $[11]$ and references therein), since the latter involve many complex physical processes, some of which might not be necessary or relevant for my own calculations. To minimize computational expenses, I will build a program from bottom-up that includes select physical and chemical processes based on the local simulations from steps 1 and 2. This will allow me to predict the \textbf{range of planet compositions that can form under reasonable disk conditions}, as well as \textbf{constrain a planet's formation location based on its chemical composition}. Comparing my results with current observations of atmospheric spectra (e.g., Figure 3 from $[4]$), and more importantly future JWST observations, will lead to great scientific strides in understanding the complex connection between protoplanetary disks and the formation, evolution and composition of exoplanets.

@@

%\textbf{MIT is the ideal place for me to pursue my postdoctoral work.} Its research facilities and academic excellence are unparalleled. I would love the opportunity to work with \textbf{Prof. Hilke Schlichting}, an expert in planet formation theory and dynamics --- such a collaboration would be instrumental in achieving my research goals during my postdoctoral tenure and afterwards. I would also like to collaborate with \textbf{Prof. Sara Seager}, a co-investigator of the TESS mission and a leading theorist in atmospheric chemistry, as well as with members of her group, who specialize in a broad range of exoplanet theory and computational topics. \textbf{Dr. Margaret Pan}, an excellent theorist in planetary dynamics and protoplanetary disks, will join the MIT EAPS department in early 2016 --- it would be a pleasure to work with her as well. Additionally, the MIT Physics department hosts leaders in detecting and characterizing worlds outside the Solar system, such as \textbf{Prof. Joshua Winn}, which presents great prospects in connecting my theoretical research work with observations. \\

@@

%\section*{References}
\footnotesize
%\begin{thebibliography} {9}
\noindent $[1]$ Batalha, N. M. 2014, Proceedings of the National Academy of Science, 111, 12647 \\
$[2]$ Lissauer J. J., Dawson R. I., Tremaine S., 2014, Nature, 513, 336 \\
$[3]$ Debes, J. H., Jang-Condell, H., Weinberger, A. J., Roberge, A., \& Schneider, G. 2013, ApJ, 771, 45 \\
$[4]$ Kreidberg, L., Line, M. R., Bean, J. L., et al. 2015, ApJ, 814, 66 \\
\textbf{$[5]$ Piso, A.-M. A., \& Youdin, A. N. 2014, ApJ, 786, 21} \\
\textbf{$[6]$ Piso, A.-M. A., Youdin, A. N., \& Murray-Clay, R. A. 2015, ApJ, 800, 82} \\
\textbf{$[7]$ Piso, A.-M. A., \"Oberg, K. I. , Birnstiel, T., \& Murray-Clay, R. A. 2015, ApJ, 815, 109} \\
\textbf{$[8]$ Piso, A.-M. A,. Pegues, J., \& \"Oberg, K. I.  2016, ApJ, 833, 203} \\
$[9]$ ALMA Partnership et al., 2015, ApJ, 808, L3 \\
$[10]$ \"Oberg, K. I., Murray-Clay, R., \& Bergin, E. A. 2011, ApJ, 743, L16 \\
$[11]$ Ida, S., Lin, D. N. C., \& Nagasawa, M. 2013, ApJ, 775, 42 \\
$[12]$ Semenov, D., \& Wiebe, D. 2011, ApJS, 196, 25
%\end{thebibliography}
